---
title: "Assignment 1, Beck"
author: "Catherine Beck"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##INPUT ALL THE DATA: 

```{r}
library(tidyverse)
library(ggplot2)


years <- 2017:2021
quarters <- 1:4
types <- c("Electric", "Gas")

pge_data <- NULL


for(year in years){
  for(quarter in quarters) {
    for(type in types){
      filename <- paste0(
          "pge/PGE_",
          year,
          "_Q",
          quarter,
          "_",
          type,
          "UsageByZip.csv"
        )
  
  ##show the file names 
  print(filename)
  
  ##skip quarters 3 and 4 of year 2021
  if((year == 2021)& (quarter %in% 3:4))next
  
  ## actually reading the data
  temp <- read_csv(filename)
  
  ##converting the units for electric column
  if(type == "Electric"){
    temp <- temp %>% 
      mutate(TOTALKBTU = TOTALKWH * 3.41214) %>%  ##conversion rate
      select(-TOTALKWH,-AVERAGEKWH)
  }
  
  ##converting the units for gas column 
  if(type == "Gas"){
    temp <- temp %>% 
      mutate(TOTALKBTU = TOTALTHM * 99.9761) %>% 
      select(-TOTALTHM, -AVERAGETHM)
  }
 
  ##bind and set the data
  pge_data <- rbind(pge_data,temp)
  saveRDS(pge_data, "pge_data.rds")

    }
  }
}
```

## Filter customer class to only have residential/commerical gas/electric (4 bodies)
```{r}
pge_filter <- 
  pge_data %>%
  filter(
    CUSTOMERCLASS %in% c("Gas- Commercial", "Elec- Commercial", "Gas- Residential", "Elec- Residential"),
    ) %>%
    group_by(
      MONTH,
      YEAR,
      CUSTOMERCLASS
    )%>%
      summarize(
        TOTALKBTU = sum(TOTALKBTU, na.rm = TRUE)
      ) %>%
      mutate(
        DATE =
          paste(
            YEAR,
            MONTH,
            "01",
            sep = "-"
          ) %>% as.Date()
      )
  
```


##Create the residential bar chart
```{r}
pge_residential_chart <-
  pge_filter %>% 
  filter(CUSTOMERCLASS %in% c("Gas- Residential", "Elec- Residential")) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = DATE %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  theme(axis.text.x = element_text(angle = 90))+ 
  labs(
    x = "Year",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electricity Usage 2017-2020",
    fill = "Electricity Type"
  )
## change the date range from monthly to yearly ==> change the x axis.
##pge_residential_chart + scale_x_continuous(breaks = c(2017, 2018, 2019, 2020, 2021))


#plot the chart
pge_residential_chart
```

```{r}
pge_commercial_chart <-
  pge_filter %>% 
  filter(CUSTOMERCLASS %in% c("Gas- Commercial", "Elec- Commercial")) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = DATE %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  theme(axis.text.x = element_text(angle = 90))+ 
  labs(
    x = "Months",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electricity Usage 2017-2020",
    fill = "Electricity Type"
  )
 
## change the date range from monthly to yearly ==> change the x axis.
##pge_residential_chart + scale_x_continuous(breaks = c(2017, 2018, 2019, 2020, 2021))


#plot the chart
pge_commercial_chart
```

##plot a third chart for 2020 electricity usage 

```{r}
pge_2020_chart <-
  pge_filter %>% 
  filter(CUSTOMERCLASS %in% c("Gas- Commercial", "Elec- Commercial", "Gas- Residential", "Elect- Commercial")) %>% 
  filter(year %in% 2020) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = DATE %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  theme(axis.text.x = element_text(angle = 90))+ 
 
  labs(
    x = "Months",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electricity Usage 2020",
    fill = "Electricity Type"
  )
## change the date range from monthly to yearly ==> change the x axis.
##pge_residential_chart + scale_x_continuous(breaks = c(2017, 2018, 2019, 2020, 2021))


#plot the chart
pge_2020_chart
```

##plot a fourt chart for 2019 electricity usage 
```{r}
pge_2019_chart <-
  pge_filter %>% 
  filter(CUSTOMERCLASS %in% c("Gas- Commercial", "Elec- Commercial", "Gas- Residential", "Elect- Commercial")) %>% 
  filter(year %in% c("2019")) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = DATE %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  theme(axis.text.x = element_text(angle = 90))+ 
  labs(
    x = "Months",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Electricity Usage 2019",
    fill = "Electricity Type"
  )
## change the date range from monthly to yearly ==> change the x axis.
##pge_residential_chart + scale_x_continuous(breaks = c(2017, 2018, 2019, 2020, 2021))


#plot the chart
pge_2019_chart
```

##Results from Comparison of 2020 vs 2019

Interestingly, the usage of all electricity/gas commercially and residentially dropped in 2020. My gut instic was that the residential energy consumption would have increased with people spending most of their time during the pandemic in their homes and working from home. By the winter months of 2020, energy usage goes back up for both residential and commercial usage (with a greater emphasis on commercial usage). 
