---
title: "Beck_Assignment2"
author: "Catherine Beck"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

Download all of the necessary packages and access Census data with key
```{r}
library(tidyverse)
library(sf)
library(tigris)
library(censusapi)
library(mapview)
library(tidycensus)
library(leaflet)
Sys.setenv(CENSUS_KEY ="c18a7f6e5cbbd1e7f0d3047d8b674cbaad89bcb3")
```


ACCESS THE 2020 POPULATION SUBDIVIDED BY RACE DATA
```{r}

##access the census data
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
  )

##Access the population data 

scc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "P1_001N"
  )%>% 
  transmute(
    block =
      paste0(state,county,tract,block),
    pop_2020 = P1_001N
  )
scc_blocks_2020 <- blocks("CA", "Santa Clara", year = 2020, progress_bar = F)

```


SAVE THE scc_blocks_2020`
```{r}
saveRDS(scc_blocks_2020, "scc_blocks_2020.rds")
scc_blocks_2020 <- readRDS("scc_blocks_2020.rds")
```


MAP THE 2020 BLOCKS DATA ALLIGNED GEOMETRY SET 
```{r}
la_boundary <- places("CA", progress_bar = F) %>% 
  filter(NAME == "Los Altos")

la_blocks_2020 <-  scc_blocks_2020 %>% 
  st_centroid() %>% 
  .[la_boundary, ] %>% 
  st_drop_geometry()  %>% 
  left_join(scc_blocks_2020 %>% select(GEOID20)) %>% ##when joining two cols you need to specify what cols to join across the two datasets
  st_as_sf() %>%
  select(GEOID20) %>%
  left_join(scc_pop_2020 %>% rename(GEOID20 = block))

##mapview(la_blocks_2020, alpha.regions = 0, color = "red", lwd = 2)
```
SAVE THE la_blocks_2020
```{r}
saveRDS(la_blocks_2020, "la_blocks_2020.rds")
la_blocks_2020 <- readRDS("la_blocks_2020.rds")
```

SPATIAL SUBSETTING 
```{r}
la_blocks_2020 <- scc_blocks_2020 %>%
  .[la_boundary, ] %>% 
  select(GEOID20) %>%
  left_join(scc_pop_2020 %>% rename(GEOID20 = block))

```



GET DECEN. POP DATA FOR 2010 
```{r}
dec_vars_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl",
    type = "variables"
  )

scc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:085",
    vars = "P001001"
  ) %>% 
  transmute(
    block =
      paste0(state,county,tract,block),
    pop_2010 = P001001
  )

scc_blocks_2010 <- blocks("CA", "Santa Clara", year = 2010, progress_bar = F)
```

SAVE THE scc_blocks_2010
```{r}
saveRDS(la_blocks_2020, "scc_blocks_2010.rds")
scc_blocks_2010 <- readRDS("scc_blocks_2010.rds")
```


MAP 2010 BLOCKS ALLIGNED GEOMETRY SET: 

```{r}
la_boundary <- places("CA", progress_bar = F) %>% 
  filter(NAME == "Los Altos")

la_blocks_2010 <-  scc_blocks_2010 %>% 
  st_centroid() %>% 
  .[la_boundary, ] %>% 
  st_drop_geometry()  %>% 
  left_join(scc_blocks_2010 %>% select(GEOID10)) %>% 
  st_as_sf() %>%
  select(GEOID10) %>%
  left_join(scc_pop_2010 %>% rename(GEOID10 = block))
```

SPATIAL SUBSETTING METHOD FROM 2.4

```{r}
la_blocks_2020_intersect <- la_blocks_2020 %>%
 st_transform(26910) %>%
  mutate(original_area = st_area(.)) %>%
  st_intersection(
   la_blocks_2010 %>%
     select(GEOID10) %>%
     st_transform(26910)
  )%>%
  mutate(
    leftover_area = st_area(.),
    perc_area = as.numeric(leftover_area / original_area),
    pop_2020 = pop_2020 * perc_area
  )
```

POPULATION CHANGE 

```{r}
la_blocks_2020_reshaped <- la_blocks_2020_intersect %>%
  st_drop_geometry() %>% ##we just want to focus on the "broken pieces" and putting them together
  group_by(GEOID10)%>% ##GEOID10 > GEOID20 but either works
  summarize(
    pop_2020 = sum(pop_2020, na.rm = T) %>% round(), ##add all the small pop counts to a larger pop count
  )
```

##turns off the previous code for knitting at the end, do this to all above
```{r, eval = F} 

```



```{r}
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs"
la_pop_2010_2020 <- la_blocks_2010 %>%
  st_transform(projection) %>%
  left_join(la_blocks_2020_reshaped) %>% ##blocks now have GEOIDs and pop info. 
  mutate(
    absolute_pop = (pop_2020 - pop_2010) / as.numeric(st_area(.)) * 43560 ##pop per acre, 43560 converts from sq ft to acres
  )
##will help knit for overflow issues 
saveRDS(la_pop_2010_2020, "la_pop_2010_2020.rds")
```

```{r}
max <- max(abs(la_pop_2010_2020$absolute_pop)) ##take every single pop change you created and makes them all abs values and tells you what the biggest one was. 
absolute_pal <- colorNumeric(
  palette = "PiYG",
  domain = c(-max, max) #feeding the data which was placed in the max variable into the symmetrical color palette. 
)
```

MAP
```{r}
la_pop_2010_2020 <- readRDS("la_pop_2010_2020.rds") ##will help knit for overflow issues 

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = la_pop_2010_2020,
    fillColor = ~absolute_pal(absolute_pop), ##use the pal relevant to your range of colors)
     color = "white",
    opacity = 0.5,
    fillOpacity = 0.75,
    weight = 1,
    label = ~paste0(
      round(absolute_pop)
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = la_pop_2010_2020,
    pal = absolute_pal,
    values = ~absolute_pop,
    title = "Population Difference in Los Altos Between 2010 and 2020"
  )
```



REFLECTION:

I think I am missing something having to do with density? Overall I found this assingment really difficult. As a beginner coder, I feel super unsuported when it comes to programming. Though I appreciate the office hours (which meet during my class so cannot attend) and the slack channel, I feel I am missing a fundamental understanding of programming. Workshops, labs, or crash courses would be super helpful. Any time you all give me help, I don't know how to interpret it because I don't know fundamentals of programming. 