---
title: "Beck_Assignment5"
author: "Catherine Beck" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(tidyverse)
library(censusapi)
library(sf)
library(mapview)
library(tigris)
library(readxl)
library(leaflet)



Sys.setenv(CENSUS_KEY="c18a7f6e5cbbd1e7f0d3047d8b674cbaad89bcb3")

ces4 <- read_excel(
    "calenviroscreen40resultsdatadictionary_F_2021.xlsx", 
  sheet = "CES4.0FINAL_results"
)
colnames(ces4)

##clean up the dataset
ces4_clean <- ces4 %>%
  select(!ends_with("Pctl"))

```

LOAD BAY PM 2.5 DATA
```{r}
bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

ca_tracts <- tracts("CA", cb = T, progress_bar = F)

ces4_bay_pm25 <-
  ces4 %>% 
  filter(`California County` %in% bay_county_names) %>% 
  select(`Census Tract`, PM2.5) %>% 
  left_join(
    ca_tracts %>% 
      transmute(GEOID = as.numeric(GEOID)), 
    by = c("Census Tract" = "GEOID")
  ) %>% 
  st_as_sf()

pm25_pal <- colorNumeric(
  palette = "Reds",
  domain = ces4_bay_pm25$PM2.5
)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = ces4_bay_pm25,
    fillColor = ~pm25_pal(PM2.5),
    color = "white",
    weight = 0.5,
    fillOpacity = 0.5,
    label = ~PM2.5
  )
```
The areas of darker red are areas with higher concentration of PM 2.5, the lighter the area, the less concentration of  PM 2.5 There is a lot of PM 2.5 located in the East Bay by Oakland, Richmond, and Vallejo and a lot in the South Bay by San Mateo, Fremont, Newark, and Mountain View. 

LOAD AND MAP ASTHMA DATA

```{r}
bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

ca_tracts <- tracts("CA", cb = T, progress_bar = F)

ces4_bay_asthma_pm25 <-
  ces4 %>% 
  filter(`California County` %in% bay_county_names) %>% 
  left_join(
    ca_tracts %>% 
      transmute(GEOID = as.numeric(GEOID)), 
    by = c("Census Tract" = "GEOID")
  ) %>% 
  st_as_sf()

asthma_pal <- colorNumeric(
  palette = "Reds",
  domain = ces4_bay_asthma_pm25$Asthma
)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = ces4_bay_asthma_pm25,
    fillColor = ~asthma_pal(Asthma),
    color = "white",
    weight = 0.5,
    fillOpacity = 0.5,
    label = ~Asthma
  )
```
The areas of darker red are areas with higher concentration of asthma, the lighter the area, the less concentration of people with asthma. There is a lot of asthma located in the East Bay by Oakland, Richmond, and Vallejo. 

Clean the Data and Create the Scatter Plot with PM2.5 on the x-axis and Asthma on the y-axis
```{r}
bay_asthma_pm_tract <-
  ces4_bay_asthma_pm25 %>%
  filter(
    !is.na(Asthma),
    !is.na(PM2.5)
  )
ggplot(
    data = bay_asthma_pm_tract,
    aes(
      x = PM2.5,
      y = Asthma
    )
  ) +
  geom_point()+
  labs(
    title = "Asthma x Pm 2.5 Scatterplot"
  ) +
  geom_smooth(method = "lm")
```
At this stage, the best fit line is not very clear with the points. But that is because there are several outliers and the points themselves are not balanced. 

Coefficients
```{r}
model <-  lm(Asthma ~ PM2.5,bay_asthma_pm_tract )

summary(model)
```

An increase of PM 2.5 by one unit appears to be  associated with an increase of Asthma by 19.862; With a p-value in the 0.01 to 0.05 range. This also conceptually makes sense and is a big part of what EJ activists are fighting about when it comes to Cap and Trade not covering local air pollution. 

Plot Residuals 
```{r}
plot(density(residuals(model)))
```
Ideally our residual plots look like a bell curve. This curve is skewed to the left and unbalanced. 

Run log on and  Create the Scatter Plot with PM2.5 on the x-axis and Asthma on the y-axis
```{r}
bay_asthma_pm_tract <-
  ces4_bay_asthma_pm25 %>%
  filter(
    !is.na(Asthma),
    !is.na(PM2.5)
  )
ggplot(
    data = bay_asthma_pm_tract,
    aes(
      x = PM2.5,
      y = log(Asthma)
    )
  ) +
  geom_point()+
  labs(
    title = "Asthma x Pm 2.5 Scatterplot"
  ) +
  geom_smooth(method = "lm")
```
Coefficients
```{r}
model <-  lm(log(Asthma) ~ log(PM2.5),bay_asthma_pm_tract )

summary(model)
plot(density(residuals(model)))
```
This distribution looks way better, though the dip at what should be the peak of the bell curve is a bit confusing. This realtionship between asthma and PM 2.5 and residuals looks more normal though. 



