---
title: "Beck_Assignment4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##Download libraries
```{r}
library(tidyverse)
library(sf)
library(tigris)
library(censusapi)
library(mapview)
library(tidycensus)
library(leaflet)
library(readxl)
Sys.setenv(CENSUS_KEY ="c18a7f6e5cbbd1e7f0d3047d8b674cbaad89bcb3")

acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )

acs_vars_2019_1yr <-
  listCensusMetadata(
    name = "2019/acs/acs1",
    type = "variables"
  )
```



##Download the data from ACS B19001 tables to choose for race/education add up a - g for pop by using a for loop
```{r}
census_race_categories <- 
  c(
    "White Alone",
    "Black or African American",
    "American Indian and Alaska Native Alone",
    "Asian Alone",
    "Native Hawaiian and Other Pacific Islander Alone",
    "Some Other Race Alone",
    "Two or More Races"
  )

bay_ed_race <-
  1:7 %>% 
  map_dfr(function(x){
    getCensus(
      name = "acs/acs5",
      vintage = 2019,
      region = "county:085",
      regionin = "state:06",
      vars = paste0("group(C15002",LETTERS[x],")")
    ) %>%
      select(!c(GEO_ID,state,county,NAME) & !ends_with(c("EA","MA","M"))) %>%
      pivot_longer(
        ends_with("E"),
        names_to = "name",
        values_to = "estimate"
      ) %>%
      left_join(
        acs_vars_2019_5yr %>% 
          select(name, label)
      ) %>% 
      select(-name) %>% 
      separate(
        label,
        into = c(NA,NA,NA,"education"),
        sep = "!!"
      ) %>% 
      group_by(education) %>% 
      summarize(across(everything(), sum)) %>% 
      filter(!is.na(education)) %>% 
      mutate(race = census_race_categories[x])
  })
```

##Graph 
```{r}
bay_race_total <-
  bay_ed_race %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(education = "Total")

bay_ed_race %>% 
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = rev(c("Total",unique(bay_ed_race$education)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_ed_race$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
     x = "Educational Attainment",
    y = "Proportion of Population 25 years or older",
    title = "San Mateo County Educational Attainment by race",
    fill = "Race"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```


Comments:

The proportion of educational attainment is overwhelmingly attributed to the white population, especially when you get to secondary education and University/beyond level. This is not suprising, but visualizing the graph and seeing how small the represetnation of Black/African-American and Indigenous populations are is very disheartening. San Mateo county is home to some of the greatest discrimination in this state, and seeing how that plays out across education levels is even more evidence to that fact. 

6.5

Separate Between Latinx and White Non Latinx

```{r}
 census_ethnicity_categories <- data.frame(
  code = c("H", "I"),
  category =
    c(
      "White Not Latinx",
      "Latinx"
    )
)

bay_ed_ethnicity <-
  1:2 %>% 
  map_dfr(function(x){
    getCensus(
      name = "acs/acs1",
      vintage = 2019,
      region = "county:085",
      regionin = "state:06",
      vars = paste0("group(C15002",census_ethnicity_categories$code[x],")")
    ) %>%
      select(!c(GEO_ID,state,county,NAME) & !ends_with(c("EA","MA","M"))) %>%
      pivot_longer(
        ends_with("E"),
        names_to = "name",
        values_to = "estimate"
      ) %>%
      left_join(
        acs_vars_2019_1yr %>% 
          select(name, label)
      ) %>% 
      select(-name) %>% 
      separate(
        label,
        into = c(NA,NA,NA,"education"),
        sep = "!!"
      ) %>% 
      filter(!is.na(education)) %>% 
      mutate(ethnicity = census_ethnicity_categories$category[x])
  })


```
Collect total number of people across all ethnicities
```{r}
bay_education_total_2019 <-
  getCensus(
      name = "acs/acs1",
      vintage = 2019,
      region = "county:085",
      regionin = "state:06",
      vars = "group(C15002)"
  ) %>%
   select(!c(GEO_ID,state,county,NAME) & !ends_with(c("EA","MA","M"))) %>%
      pivot_longer(
        ends_with("E"),
        names_to = "name",
        values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_1yr %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
      into = c(NA,NA,NA,"education"),
      sep = "!!"
  ) %>% 
  filter(!is.na(education)) %>% 
  mutate(
    ethnicity = "Total",
    education = case_when(
      education %in% c("Less than 9th grade", "9th to 12th grade, no diploma") ~ "Less than high school diploma",
      education == "High school graduate (includes equivalency)" ~ education,
      education %in% c("Some college, no degree", "Associate's degree") ~ "Some college or associate's degree",
      education %in% c("Bachelor's degree", "Graduate or professional degree") ~ "Bachelor's degree or higher"
    )
  ) %>%
  group_by(education, ethnicity) %>% 
  summarize_all(sum) ##syntax method equiv. to above

##Doing all of the above to the Latinx/White not Latinx Dataset 
bay_education_ethnicity_2019 <- 
  bay_ed_ethnicity %>%
    group_by(education, ethnicity) %>% 
    summarize_all(sum) %>% 
    rbind(bay_education_total_2019) %>% 
    pivot_wider(
      names_from = "ethnicity",
      values_from = "estimate"
    ) %>% 
    mutate(
      `Not White, not Latinx` = Total - Latinx - `White Not Latinx` 
    ) %>% 
    select(-Total) %>% 
    pivot_longer(
      -c(education), ##left out of original 
      names_to = "ethnicity",
      values_to = "estimate"
    )
```



Graph:

```{r}
bay_education_ethnicity_2019_stacked <- ##Lets you add a total ethnicity 
  bay_education_ethnicity_2019 %>%
  rbind(bay_education_ethnicity_2019
      %>%
      group_by(ethnicity) %>% 
      summarize(estimate = sum(estimate)) %>% 
      mutate(education = "Total")
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = rev(c("Total", unique(bay_education_ethnicity_2019$education)))),
      y = estimate,
      fill = ethnicity %>% factor(levels = rev(unique(bay_education_ethnicity_2019$ethnicity)))
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Educational attainment",
    y = "Population 25 years or older",
    title = "Bay Area educational attainment by ethnicity",
    fill = "Ethnicity"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  ) +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
bay_education_ethnicity_2019_fill <-
  bay_education_ethnicity_2019
bay_education_ethnicity_2019_stacked
bay_education_ethnicity_2019_fill
```



Commentary: 
Once again, there is an overrepresentation of White, not latinx folks within  educational attainment in the Bay. There seems to be a nearly equal amount of white, not latinx people as the population of not white, not latinx which is a huge overrepresentation of one ethnicity/race in an educational setting. 